---
title: "CFAR Social Network Project"
output: pdf_document
---
```{r setup, include=FALSE}
library(formatR) #
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 50), tidy = TRUE) #
```
**<u>PI</u>**: Dr. Jeffrey Campbell <br>
**<u>Author</u>**: Max Schultz 

# Preliminaries
```{r echo = T, results = "hide", message = F, warnings = F}
library(tidyr) # data cleaning processing 
library(dplyr) # further data manipulation (%>% (pipe) operator)
library(igraph) # social networks 
library(RColorBrewer) # color palettes 
library(stringr) # regex 
library(purrr) # used to create the network objects 
```


# Data Loading and Cleaning
Raw data from RedCap was given in CSV format. Reading data in and cleaning:

```{r}
df_messy = read.csv("/Users/maxschultz/Desktop/CFAR Social Network Project/CHERISHSN_DATA_2025-10-16_1847.csv") 
```

The raw RedCap data is in wide format, where each row corresponds to an index patient. Contact information is stored in these wide columns. 
We want a row for each contact for our analysis; thus, I use the **pivot_longer()** function from the tidyr library to convert wide data into long data. The columns that are currently useful for us are the following: 

**Contact-level attributes:**

- contact id (primary key)
- contact age
- contact sex
- contact relation to child index
- primary caregiver 
- household id
- previous tb 
- contact information (lived, stayed, visited, occasional, elsewhere only)
- other contact information (same room, same bed)
- contact tb symptoms (weight loss, cough)
- contact tb risk elements (drug use, smoking)

**Index-level attributes:**

- index ID (not unique because of long format)
- index age 
- index sex
- final IGRA result 
- study group (CHEU vs. CHUU)

Contact data was collected over time, and spans numbers 1-28. For example, household id is stored as hh_id_1cf and goes all the way to hh_id_28cf and this is the case for all contact variables.
To mitigate this, we'll use the **pivot_longer()** function with regular expressions to capture all of these columns. 

```{r}
df_cleaned <- df_messy %>%
    pivot_longer( # these need only be the contact-level variables
        cols = matches(
            "^(snml_sex|snml_vage|snml_relate|hh_id|pcare|tb_dx|lived|stayed|visited|occas|elswe|room|bed|smoke|drug)_\\d+cf"
          ),
        names_to = c(".value", "ctc_idx"),
        names_pattern = "^(snml_sex|snml_vage|snml_relate|hh_id|pcare|tb_dx|lived|stayed|visited|occas|elswe|room|bed|smoke|drug)_(\\d+)cf(?:___1)?$",
        values_drop_na = FALSE
    ) %>% # end pivot longer scope
    mutate( # creating contact_id as pervious forms have used contact x from index y; i.e. ctc_x_idx_y
    ctc_idx = as.integer(ctc_idx),
    contact_id = str_c("ctc_", record_id, "_idx_", ctc_idx)
    ) %>%
    filter(!( # failsafe filtering just in case the regular expression is picky 
        is.na(snml_sex) & 
        is.na(snml_vage) & 
        is.na(snml_relate) & 
        is.na(hh_id))) %>%
    arrange(record_id, ctc_idx) %>%
    transmute( # renaming and addressing data types 
    record_id = as.character(record_id),
    contact_id = as.character(contact_id),
    enrol_age_mrg = as.numeric(enrol_age),
    enrol_age_yrs_mrg = as.integer(floor(enrol_age_mrg/365.25)),
    cherish_sn_study_group_mrg = factor(cherish_sn_study_group),
    final_igra_results = factor(igra_results_cfar_complete),
    snml_vage_mrg = as.numeric(snml_vage),
    snml_sex_mrg = factor(snml_sex),
    snml_relate_mrg = factor(snml_relate),
    hh_id_unified_ = factor(hh_id),      
    pcare_mrg = factor(pcare),
    smoke_mrg = factor(smoke),
    drug_mrg = factor(drug),
    tb_dx_mrg = factor(tb_dx),
    lived_mrg = factor(lived),
    stayed_mrg = factor(stayed),
    visited_mrg = factor(visited),
    occas_mrg = factor(occas),
    elswe_mrg = factor(elswe),
    room_mrg = factor(room),
    bed_mrg = factor(bed),
  )
head(df_cleaned)
names(df_cleaned)
```

This is a large code chunk, but essentially what I did was remove the various sparse columns and truncate them into one row per contact. 

# Social Network Plots: Helper Functions 
We will be using two different types of social network plots in our analysis: a multi-ego network, and a multi-ego network with added edges corresponding to a shared household.
Let's build these graph objects using **igraph**. 
```{r}
network <- function(data, type = c("basic", "intra_household")){
    type <- match.arg(type)

    # data columns needed to build the graphs 
    index <- data$record_id
    hh <- data$hh_id_unified_

    # base attributes 
    ego_basic <- data %>%
      transmute(from = record_id, to = contact_id) %>%
      filter(!is.na(from), !is.na(to), from != to)


    # basic multi-ego network
    if (type == "basic"){ 
        edges <- ego_basic
    }


    # multi-ego network with intra household edges 
    # if (type == "intra_household" && "hh_id_unified_" %in% names(data) && any(!is.na(hh))) {
    #     hh_edges <- data %>%
    #         filter(!is.na(hh), !is.na(contacts)) %>% # keeping rows where both of these are not null
    #         group_by(record_id, hh_id_unified_) %>%  # and further stratifying by index and household (this is our ego level information for our graphs)
    #         summarise(
    #           pairs = list({ # this is just a matrix of all the unordered edges depending on what is in the group by clause
    #             u = unique(contacts)
    #             if (length(u)<2){
    #               return(NULL)
    #             } else{
    #               m = t(combn(u,2)) # all combinations of edges for two contacts 
    #               tibble::tibble(from = m[, 1], to = m[, 2])
    #             }
    #           }),
    #           .groups = "drop" # just removing any potential groupings so its just a tibble 
    #         ) %>%
    #         tidyr::unnest(pairs) %>% # converting to cols in the df 

    #     edges <- bind_rows(ego_basic, hh_edges) %>% distinct() # now this adds the information from the original graph back into this one and removes any duplicates
    # }
    if (type == "intra_household" && "hh_id_unified_" %in% names(data) && any(!is.na(hh))) {

      hh_edges <- data %>%
        dplyr::filter(!is.na(hh_id_unified_), !is.na(contact_id)) %>%
        dplyr::group_by(record_id, hh_id_unified_) %>%
        dplyr::summarise(
          pairs = list({
            u <- unique(contact_id)
            if (length(u) < 2) {
              tibble::tibble(from = character(0), to = character(0))
            } else {
              m <- t(combn(u, 2))
              tibble::tibble(from = m[, 1], to = m[, 2])
            }
          }),
          .groups = "drop"
        ) %>%
        tidyr::unnest(pairs)

      edges <- dplyr::bind_rows(ego_basic, hh_edges) %>% dplyr::distinct()
    }

    edges <- edges %>%
        mutate(a = pmin(from, to), b = pmax(from,to))%>% # I am mapping edges to form undirected unique edges 
        distinct(a,b) %>%
        transmute(from = a, to = b)

    graph = graph_from_data_frame(edges, directed = FALSE)
    egos <- unique(as.character(index))
    V(graph)$type <- ifelse(V(graph)$name %in% egos, "index", "contact")
    V(graph)$size <- ifelse(V(graph)$type == "index", 3, 2) # this is just a sizing convention i've found works well with the plots 
    return(graph)
}
```

Now that the graph objects have been made, lets create some graph and coloring conventions for the different plots we want.

```{r}
alpha = 0.5

global_code = list(
  alpha = alpha, 
  contact_age_levels = 1:10,
  contact_age_labels = c("0-2y","3-5y","7-11y","12-15y","16-20y","21-30y","31-40y","41-50y","51-60y",">60y"),
  palettes = list( # all colors can be changed according to needs. See base R colors here: https://www.datanovia.com/en/blog/awesome-list-of-657-r-color-names/
      basic_network = c(index = "red", contact = "lightblue"),
      contact_age_network = setNames(RColorBrewer::brewer.pal(10, "Spectral"),
        c("0-2y", "3-5y", "7-11y", "12-15y", "16-20y", "21-30y", "31-40y", "41-50y", "51-60y", ">60y")),
      igra_network = c("2" = "red", "1" = "purple", "0" = "black"),
      hiv_network = c("2" = "red", "1" = "purple"),
      hh_info_network = c(lived = "indianred", stayed = "darkorange1", occas = "goldenrod1",
              elswe = "chartreuse4", visited = "blue"),
      hh_id_network = c(
        "1"="brown1","2"="chocolate1","3"="yellow","4"="aquamarine3",
        "5"="cornflowerblue","6"="darkorchid2","7"="deeppink3",
        "NA"=rgb(0.5,0.5,0.5,alpha))
  )
)

```

Now we can apply those schemes for each type of network. Check out the "color_schemes.R" file containing the full mapping code. After sourcing the file, we can instantiate a social network like so:

```{r}
source("/Users/maxschultz/Desktop/color_schemes.r")
```

## Basic Multi-Ego Network 

```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)
basic = network(data = df_cleaned, type = "basic")
basic_graph = color_schemes(graph = basic, data = df_cleaned, global_code = global_code, scheme = "basic")
plot.igraph(
  basic_graph$graph,
  layout = layout_with_fr(basic_graph$graph),
  vertex.color = V(basic_graph$graph)$plot_color,
  vertex.size  = V(basic_graph$graph)$size,
  vertex.label = NA,
  edge.color   = "gray70",
  edge.width   = 0.5,
  main = "Basic Multi-Ego Network")
legend(
  "right",
  legend = basic_graph$legend$labels,
  #inset = c(-.35,0),
  col = basic_graph$legend$colors,
  pch = 19, # circle in R
  #pt.cex = 1.2, 
  bty = "n", 
  cex = 1
)
```


## Contact Age Network 

```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)
contact_age = color_schemes(graph = basic, data = df_cleaned, global_code = global_code, scheme = "contact_age")
plot.igraph(
  contact_age$graph,
  layout = layout_with_fr(contact_age$graph),
  vertex.color = V(contact_age$graph)$plot_color,
  vertex.size  = V(contact_age$graph)$size,
  vertex.label = NA,
  edge.color   = "gray70",
  edge.width   = 0.5,
  main = "Contact Age Multi-Ego Network"
)
legend(
  "right",
  legend = contact_age$legend$labels,
  #inset = c(-.35,0),
  col = contact_age$legend$colors,
  pch = 19, # circle in R
  #pt.cex = 1.2, 
  bty = "n", 
  cex = 1
)
```

## Index HIV and Contact Age Network (Combined)
This next plot is rather interesting, because the way we have defined these structures allows us to stitch two schemes together... with the exception of the legends.  
```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)

# network definitions
index_hiv = color_schemes(graph = basic, data = df_cleaned, global_code = global_code, scheme = "hiv_exposure") # Index HIV network 
contact_age_scheme = color_schemes(graph = basic, data = df_cleaned, global_code = global_code, scheme = "contact_age") # Contact Age Network 
graph_hiv = index_hiv$graph # just the graph 
contact_age_palette = global_code$palettes$contact_age_network
label_by_level = setNames(global_code$contact_age_labels, global_code$contact_age_levels) # age mapping 

# you can use the legends provided and override, or just create a new legend.. whatever is easiest 
# overriding 
contact_age = df_cleaned$snml_vage_mrg
names(contact_age) = df_cleaned$contact_id # contact age lookup table 
contacts = which(V(graph_hiv)$type == "contact") 
level = contact_age[V(graph_hiv)$name[contacts]] # nodes that are contacts 
ok = !is.na(level)
V(graph_hiv)$plot_color[contacts[ok]] <- unname(contact_age_palette[label_by_level[as.character(level[ok])]]) # mapping colors to non-null contact age
V(graph_hiv)$plot_color[contacts[!ok]] <- "white" 

plot.igraph(
  graph_hiv,
  layout = layout_with_fr(graph_hiv),
  vertex.color = V(graph_hiv)$plot_color,
  vertex.size  = V(graph_hiv)$size,
  vertex.label = NA,
  edge.color   = "gray70",
  edge.width   = 0.5,
  main = "Contact Age & Index HIV Multi-Ego Network"
)

legend(
  "topright",
  legend = contact_age_scheme$legend$labels,
  #inset = c(-.35,0),
  col = contact_age_scheme$legend$colors,
  pch = 19, # circle in R
  #pt.cex = 1.2, 
  bty = "n", 
  cex = 1
)
hiv_leg <- index_hiv$legend
keep <- grepl("^Index", hiv_leg$labels) # keeping only index HIV entries
hiv_leg$labels <- hiv_leg$labels[keep]
hiv_leg$colors <- hiv_leg$colors[keep]
legend(
  "right",
  legend = hiv_leg$labels,
  col    = hiv_leg$colors,
  pch    = 19,
  bty    = "n",
  cex    = 1
)
```


## Household ID Shading - Multi-Ego Network with Intra-Household Connections
```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)

intra_household <- network(df_cleaned, "intra_household") # graph initialization 

age_on_intra <- color_schemes(graph = intra_household, data = df_cleaned, global_code = global_code, scheme = "contact_age") # contact age

hh_groups <- color_schemes(graph = age_on_intra$graph, data = df_cleaned, global_code = global_code, scheme = "hh_id") # hh_id on same graph as above

plot.igraph(
  age_on_intra$graph,
  layout = layout_with_fr(age_on_intra$graph),
  vertex.color = igraph::V(age_on_intra$graph)$plot_color,
  vertex.size  = igraph::V(age_on_intra$graph)$size,
  vertex.label = NA,
  edge.color   = "gray30",
  edge.width   = 0.5,
  mark.groups  = hh_groups$groups,
  mark.col     = hh_groups$cols,     # <-- cols, not colors
  mark.border  = NA,
  mark.expand  = 5,
  main = "Contact Age + HH Shading (Intra-household edges)"
)

legend(
  "topright",
  legend = age_on_intra$legend$labels,
  col    = age_on_intra$legend$colors,
  pch    = 19,
  bty    = "n",
  cex    = 1
)
hh_leg <- hh_groups$legend
keep <- !(hh_leg$labels %in% c("Index", "Contact"))
hh_leg$labels <- hh_leg$labels[keep]
hh_leg$colors <- hh_leg$colors[keep]

legend(
  "right",
  legend = hh_leg$labels,
  col    = hh_leg$colors,
  pch    = 15,
  bty    = "n",
  cex    = 1
)
```

## Household Information Network 

```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)
hh_info = color_schemes(data = df_cleaned, global_code = global_code, graph = intra_household, scheme = "household_info")
plot.igraph(
  hh_info$graph,
  layout = layout_with_fr(hh_info$graph),
  vertex.color = igraph::V(age_on_intra$graph)$plot_color,
  vertex.size  = igraph::V(age_on_intra$graph)$size,
  vertex.label = NA,
  edge.color   = "gray30",
  edge.width   = 0.5,
)
legend(
  "right",
  legend = hh_info$legend$labels,
  col = hh_info$legend$colors,
  pch = 19,
  bty = "n",
  cex = 1
)
```

# Intra-Household Age Network with Shading for HH information

```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)
age_on_intra <- color_schemes(graph = intra_household, data = df_cleaned, global_code = global_code, scheme = "contact_age") # contact age

hh_groups <- color_schemes(graph = age_on_intra$graph, data = df_cleaned, global_code = global_code, scheme = "household_info_shading") # hh_id on same graph as above

plot.igraph(
  age_on_intra$graph,
  layout = layout_with_fr(age_on_intra$graph),
  vertex.color = igraph::V(age_on_intra$graph)$plot_color,
  vertex.size  = igraph::V(age_on_intra$graph)$size,
  vertex.label = NA,
  edge.color   = "gray30",
  edge.width   = 0.5,
  mark.groups  = hh_groups$groups,
  mark.col     = hh_groups$cols,     # <-- cols, not colors
  mark.border  = NA,
  mark.expand  = 5,
  main = "Contact Age + HH Info. Shading (Intra-household edges)"
)

legend(
  "topright",
  legend = age_on_intra$legend$labels,
  col    = age_on_intra$legend$colors,
  pch    = 19,
  bty    = "n",
  cex    = 1
)

# Example of how to just create a new legend 

legend(
  "right",
  legend = c("Lived", "Stayed", "Occas.", "Elswe.", "Visited"),
  col = c("indianred", "darkorange1", "goldenrod1", "chartreuse4", "blue"),
  pch = 15, 
  bty = "n",
  cex = 1
)

```



```{r, fig.width=11, fig.height=9, out.width="100%", fig.retina=2}
set.seed(123)
hh_info = color_schemes(data = df_cleaned, global_code = global_code, graph = intra_household, scheme = "household_info")
hh_groups <- color_schemes(graph = hh_info$graph, data = df_cleaned, global_code = global_code, scheme = "hh_id")
plot.igraph(
  hh_info$graph,
  layout = layout_with_fr(hh_info$graph),
  vertex.color = igraph::V(hh_info$graph)$plot_color,
  vertex.size  = igraph::V(hh_info$graph)$size,
  vertex.label = NA,
  edge.color   = "gray30",
  edge.width   = 0.5,
  mark.groups  = hh_groups$groups,
  mark.col     = hh_groups$cols,     # <-- cols, not colors
  mark.border  = NA,
  mark.expand  = 5,
  main = "Time Interacting (Nodes) and HH ID (Shading)"
)
legend(
  "right",
  legend = hh_info$legend$labels,
  col = hh_info$legend$colors,
  pch = 19,
  bty = "n",
  cex = 1
)
hh_leg <- hh_groups$legend
keep <- !(hh_leg$labels %in% c("Index", "Contact"))
hh_leg$labels <- hh_leg$labels[keep]
hh_leg$colors <- hh_leg$colors[keep]

legend(
  "topright",
  legend = hh_leg$labels,
  col    = hh_leg$colors,
  pch    = 15,
  bty    = "n",
  cex    = 1
)
```