---
title: "CFAR Analysis"
author: "Max Schultz"
date: "2026-06-29"
output: html_document
---

# **CFAR Analysis Code** 

### **EDA and Preprocessing** 

```{r}
library(haven)
#setwd("/Users/maxschultz/Desktop/CFAR Social Network Project")
data <- read_dta("CFAR SN 03DEC2025.dta")
cat("Data Dimensions:",dim(data),"\n") 
cat("Number of Contacts:", nrow(data))
head(data)
```

```{r clean-contact-summary-table, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(gt)

contact_data <- data

# ------------------------------------------------------------
# Helper functions
# ------------------------------------------------------------

to_num <- function(x) {
  suppressWarnings(as.numeric(as.character(x)))
}

summarize_binary <- function(data, var, label, group) {
  x <- to_num(data[[var]])
  x_binary <- ifelse(is.na(x), NA, ifelse(x == 1, 1, 0))
  
  n_total <- length(x_binary)
  n_nonmiss <- sum(!is.na(x_binary))
  n_missing <- sum(is.na(x_binary))
  n_yes <- sum(x_binary == 1, na.rm = TRUE)
  pct_yes <- round(100 * n_yes / n_nonmiss, 1)
  
  data.frame(
    Section = group,
    Variable = label,
    N = n_nonmiss,
    Missing = n_missing,
    Mean = round(mean(x_binary, na.rm = TRUE), 2),
    SD = round(sd(x_binary, na.rm = TRUE), 2),
    Median = round(median(x_binary, na.rm = TRUE), 2),
    Q1 = round(quantile(x_binary, 0.25, na.rm = TRUE), 2),
    Q3 = round(quantile(x_binary, 0.75, na.rm = TRUE), 2),
    Min = round(min(x_binary, na.rm = TRUE), 2),
    Max = round(max(x_binary, na.rm = TRUE), 2),
    Interpretation = paste0(n_yes, " yes (", pct_yes, "%)")
  )
}

summarize_numeric <- function(data, var, label, group) {
  x <- to_num(data[[var]])
  
  n_nonmiss <- sum(!is.na(x))
  n_missing <- sum(is.na(x))
  
  data.frame(
    Section = group,
    Variable = label,
    N = n_nonmiss,
    Missing = n_missing,
    Mean = round(mean(x, na.rm = TRUE), 2),
    SD = round(sd(x, na.rm = TRUE), 2),
    Median = round(median(x, na.rm = TRUE), 2),
    Q1 = round(quantile(x, 0.25, na.rm = TRUE), 2),
    Q3 = round(quantile(x, 0.75, na.rm = TRUE), 2),
    Min = round(min(x, na.rm = TRUE), 2),
    Max = round(max(x, na.rm = TRUE), 2),
    Interpretation = ""
  )
}

summarize_categorical <- function(data, var, label, group) {
  x <- data[[var]]
  tab <- table(x, useNA = "ifany")
  pct <- round(prop.table(tab) * 100, 1)
  
  most_common <- names(tab)[which.max(tab)]
  most_common_n <- max(tab)
  most_common_pct <- pct[which.max(tab)]
  
  data.frame(
    Section = group,
    Variable = label,
    N = length(x),
    Missing = sum(is.na(x)),
    Mean = NA,
    SD = NA,
    Median = NA,
    Q1 = NA,
    Q3 = NA,
    Min = NA,
    Max = NA,
    Interpretation = paste0(
      "Most common: ", most_common,
      " (", most_common_n, ", ", most_common_pct, "%)"
    )
  )
}

# ------------------------------------------------------------
# Main clean summary table
# ------------------------------------------------------------

contact_summary_clean <- bind_rows(
  
  data.frame(
    Section = "Identifier",
    Variable = "Contact ID",
    N = nrow(contact_data),
    Missing = sum(is.na(contact_data$contact_id)),
    Mean = NA,
    SD = NA,
    Median = NA,
    Q1 = NA,
    Q3 = NA,
    Min = NA,
    Max = NA,
    Interpretation = paste0(n_distinct(contact_data$contact_id), " unique contacts")
  ),
  
  summarize_numeric(
    contact_data, "snml_vage_mrg",
    "Contact age group",
    "Demographics"
  ),
  
  summarize_categorical(
    contact_data, "snml_sex_mrg",
    "Contact sex",
    "Demographics"
  ),
  
  summarize_categorical(
    contact_data, "snml_relate_mrg",
    "Relationship to index child",
    "Demographics"
  ),
  
  summarize_binary(
    contact_data, "pcare_mrg",
    "Primary caregiver",
    "Household / caregiving"
  ),
  
  summarize_categorical(
    contact_data, "hh_id_unified_",
    "Household ID",
    "Household / caregiving"
  ),
  
  summarize_binary(
    contact_data, "lived_mrg",
    "Lived with index child",
    "Contact type"
  ),
  
  summarize_binary(
    contact_data, "stayed_mrg",
    "Stayed with index child",
    "Contact type"
  ),
  
  summarize_binary(
    contact_data, "visited_mrg",
    "Visited index child",
    "Contact type"
  ),
  
  summarize_binary(
    contact_data, "occas_mrg",
    "Occasional contact",
    "Contact type"
  ),
  
  summarize_binary(
    contact_data, "elswe_mrg",
    "Elsewhere-only contact",
    "Contact type"
  ),
  
  summarize_binary(
    contact_data, "room_mrg",
    "Same room",
    "Other contact information"
  ),
  
  summarize_binary(
    contact_data, "bed_mrg",
    "Same bed",
    "Other contact information"
  ),
  
  summarize_binary(
    contact_data, "tb_dx_mrg",
    "Previous TB",
    "TB history / symptoms"
  ),
  
  summarize_binary(
    contact_data, "wloss_mrg",
    "Weight loss",
    "TB history / symptoms"
  ),
  
  summarize_binary(
    contact_data, "cough_mrg",
    "Cough",
    "TB history / symptoms"
  ),
  
  summarize_binary(
    contact_data, "smoke_mrg",
    "Smoking",
    "TB risk elements"
  ),
  
  summarize_binary(
    contact_data, "drug_mrg",
    "Drug use",
    "TB risk elements"
  )
)

# ------------------------------------------------------------
# Pretty GT table
# ------------------------------------------------------------

contact_summary_clean %>%
  gt(groupname_col = "Section") %>%
  tab_header(
    title = md("**Contact-Level Summary Statistics**"),
    subtitle = "Summary of contact demographics, household information, TB history, symptoms, and risk factors"
  ) %>%
  cols_label(
    Variable = "Variable",
    N = "N",
    Missing = "Missing",
    Mean = "Mean",
    SD = "SD",
    Median = "Median",
    Q1 = "Q1",
    Q3 = "Q3",
    Min = "Min",
    Max = "Max",
    Interpretation = "Readable summary"
  ) %>%
  fmt_number(
    columns = c(Mean, SD, Median, Q1, Q3, Min, Max),
    decimals = 2
  ) %>%
  sub_missing(
    columns = everything(),
    missing_text = ""
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
  ) %>%
  tab_options(
    table.font.size = px(13),
    data_row.padding = px(6),
    heading.title.font.size = px(18),
    heading.subtitle.font.size = px(12),
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    table.width = pct(100)
  )
```

```{r}
library(dplyr)
contact_data <- data %>%
    group_by(record_id) %>%
    summarise(num_contacts = n_distinct(contact_id) )
sum(contact_data$num_contacts)
```


```{r}
names(data)
unique(data$pcare_mrg)
length(unique(data$record_id))
```


```{r}
library(dplyr)
data$enrol_age_yrs_mrg = floor(data$enrol_age_mrg/365.25)

# data is in long format so need to avoid duplicates
index_data <- data %>% # can continue to add to this for index data 
    distinct(record_id, enrol_age_yrs_mrg)%>%
    count(enrol_age_yrs_mrg, name = "n_index")
print(index_data)

cat("Number of Index Cases:", sum(index_data$n_index))
```


```{r}
# basic contact age
contact_age_levels <- 1:10
contact_age_labels <- c("0-2y", "3-5y", "7-11y", "12-15y", "16-20y", "21-30y", "31-40y", "41-50y", "51-60y", ">60y")

contact_age <- function(data, levels = contact_age_levels, labels = contact_age_labels, na_label = "NA"){
    factor_version = factor(data, levels = levels, labels = labels)
    factor_version = addNA(factor_version)
    lev = levels(factor_version)
    lev[is.na(lev) | lev == "<NA>"] <- na_label
    levels(factor_version) <- lev
    factor_version
}
```


### **Basic Plots - Histograms** 


```{r}
# basic contact age
plot(contact_age(data$snml_vage_mrg),
    main = "Contact Age Distribution",
    ylab = "Count",
    las = 2,
    col = "lightblue")
table(contact_age(data$snml_vage_mrg))



### CONTACT AGE BY SEX
plot(contact_age(data$snml_vage_mrg[which(data$snml_sex_mrg == 1)]),
    main = "Contact Age by Sex (Male)",
    ylab = "Count",
    las = 2,
    col = "lightblue"
)
table(contact_age(data$snml_vage_mrg[which(data$snml_sex_mrg == 1)]))


plot(contact_age(data$snml_vage_mrg[which(data$snml_sex_mrg == 2)]),
    main = "Contact Age by Sex (Female)",
    ylab = "Count",
    las = 2,
    col = "lightpink"
)
table(contact_age(data$snml_vage_mrg[which(data$snml_sex_mrg == 2)]))



### CONTACT AGE BY INDEX AGE 
plot(contact_age(data$snml_vage_mrg[which(data$enrol_age_yrs_mrg == 0)]),
    main = "Contact Age by 0yo Index",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$enrol_age_yrs_mrg == 0)]))

plot(contact_age(data$snml_vage_mrg[which(data$enrol_age_yrs_mrg == 1)]),
    main = "Contact Age by 1yo Index",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$enrol_age_yrs_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$enrol_age_yrs_mrg == 2)]),
    main = "Contact Age by 2yo Index",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$enrol_age_yrs_mrg == 2)]))



### CONTACT AGE BY HIV EXPOSURE
plot(contact_age(data$snml_vage_mrg[which(data$cherish_sn_study_group_mrg == 1)]),
    main = "Contact Age by Index HIV (CHEU)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$cherish_sn_study_group_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$cherish_sn_study_group_mrg == 2)]),
    main = "Contact Age by Index HIV (CHUU)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$cherish_sn_study_group_mrg == 2)]))


### CONTACT AGE BY INDEX IGRA RESULT
plot(contact_age(data$snml_vage_mrg[which(data$final_igra_result == 0)]),
    main = "Contact Age by Index IGRA (Incomplete)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$final_igra_result == 0)]))

plot(contact_age(data$snml_vage_mrg[which(data$final_igra_result == 1)]),
    main = "Contact Age by Index IGRA (Unverified)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$final_igra_result == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$final_igra_result == 2)]),
    main = "Contact Age by Index IGRA (Complete)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$final_igra_result == 2)]))

plot(contact_age(data$snml_vage_mrg[which(is.na(data$final_igra_result))]),
    main = "Contact Age by Index IGRA (NA)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(is.na(data$final_igra_result))]))


### CONTACT AGE BY TB SYMPTOM 

# weight loss
plot(contact_age(data$snml_vage_mrg[which(data$wloss_mrg== 1)]),
    main = "Contact Age by Weight Loss (Yes)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$wloss_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$wloss_mrg== 2)]),
    main = "Contact Age by Weight Loss (No)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$wloss_mrg == 2)]))

plot(contact_age(data$snml_vage_mrg[which(data$wloss_mrg== 3)]),
    main = "Contact Age by Weight Loss (Don't Know)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$wloss_mrg == 3)]))

plot(contact_age(data$snml_vage_mrg[which(is.na(data$wloss_mrg))]),
    main = "Contact Age by Weight Loss (NA)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(is.na(data$wloss_mrg))]))

# cough 
plot(contact_age(data$snml_vage_mrg[which(data$cough_mrg== 1)]),
    main = "Contact Age by Cough (Yes)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$cough_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$cough_mrg== 2)]),
    main = "Contact Age by Cough (No)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$cough_mrg == 2)]))

plot(contact_age(data$snml_vage_mrg[which(data$cough_mrg== 3)]),
    main = "Contact Age by cough (Don't Know)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$cough_mrg == 3)]))

plot(contact_age(data$snml_vage_mrg[which(is.na(data$cough_mrg))]),
    main = "Contact Age by Cough (NA)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(is.na(data$cough_mrg))]))



### CONTACT AGE BY TB HISTORY 
plot(contact_age(data$snml_vage_mrg[which(data$tb_dx_mrg == 1)]),
    main = "Contact Age by TB History (Yes)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$tb_dx_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$tb_dx_mrg == 2)]),
    main = "Contact Age by TB History (No)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$tb_dx_mrg == 2)]))

plot(contact_age(data$snml_vage_mrg[which(data$tb_dx_mrg == 3)]),
    main = "Contact Age by TB History (Don't Know)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$tb_dx_mrg == 3)]))

plot(contact_age(data$snml_vage_mrg[which(is.na(data$tb_dx_mrg))]),
    main = "Contact Age by TB History (NA)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(is.na(data$tb_dx_mrg))]))



### CONTACT AGE BY TB RISK FACTOR 

# smoking 
plot(contact_age(data$snml_vage_mrg[which(data$smoke_mrg == 1)]),
    main = "Contact Age by Smoking (Yes)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$smoke_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$smoke_mrg == 2)]),
    main = "Contact Age by Smoking (No)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$smoke_mrg == 2)]))

plot(contact_age(data$snml_vage_mrg[which(data$smoke_mrg == 3)]),
    main = "Contact Age by Smoking (Don't Know)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$smoke_mrg == 3)]))

plot(contact_age(data$snml_vage_mrg[which(is.na(data$smoke_mrg))]),
    main = "Contact Age by Smoking (NA)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(is.na(data$smoke_mrg))]))



# drugs
plot(contact_age(data$snml_vage_mrg[which(data$drug_mrg == 1)]),
    main = "Contact Age by Drug Use (Yes)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$drug_mrg == 1)]))

plot(contact_age(data$snml_vage_mrg[which(data$drug_mrg == 2)]),
    main = "Contact Age by Drug Use (No)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$drug_mrg == 2)]))

plot(contact_age(data$snml_vage_mrg[which(data$drug_mrg == 3)]),
    main = "Contact Age by Drug Use (Don't Know)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(data$drug_mrg == 3)]))

plot(contact_age(data$snml_vage_mrg[which(is.na(data$drug_mrg))]),
    main = "Contact Age by Drug Use (NA)",
    ylab = "Count",
    las = 2
)
table(contact_age(data$snml_vage_mrg[which(is.na(data$drug_mrg))]))

```


### **Basic Plots - Violin Plots** 

```{r}
library(vioplot)
violin_contact_age <- function(x, main, col = "lightgray") {
  x <- x[!is.na(x)]                 
  vioplot(x, main = main, col = col, yaxt = "n")
  axis(2, at = contact_age_levels, labels = contact_age_labels, las = 2)
}
```

```{r}
violin_contact_age(data$snml_vage_mrg, "Contact Age Distribution", col = 'lightblue')

violin_contact_age(data$snml_vage_mrg[data$snml_sex_mrg == 1], "Contact Age By Sex (Male)", col = "lightblue")
violin_contact_age(data$snml_vage_mrg[data$snml_sex_mrg == 2], "Contact Age by Sex (Female)", col = "lightpink")

violin_contact_age(data$snml_vage_mrg[data$enrol_age_yrs_mrg == 0], "Contact Age by 0yo Index")
violin_contact_age(data$snml_vage_mrg[data$enrol_age_yrs_mrg == 1], "Contact Age by 1yo Index")
violin_contact_age(data$snml_vage_mrg[data$enrol_age_yrs_mrg == 2], "Contact Age by 2yo Index")

violin_contact_age(data$snml_vage_mrg[data$cherish_sn_study_group_mrg == 1], "Contact Age by Index HIV (CHEU)")
violin_contact_age(data$snml_vage_mrg[data$cherish_sn_study_group_mrg == 2], "Contact Age by Index HIV (CHUU)")

violin_contact_age(data$snml_vage_mrg[data$final_igra_result == 0], "Contact Age by Index IGRA (Incomplete)")
violin_contact_age(data$snml_vage_mrg[data$final_igra_result == 1], "Contact Age by Index IGRA (Unverified)")
violin_contact_age(data$snml_vage_mrg[data$final_igra_result == 2], "Contact Age by Index IGRA (Complete)")
violin_contact_age(data$snml_vage_mrg[is.na(data$final_igra_result)], "Contact Age by Index IGRA (NA)")

violin_contact_age(data$snml_vage_mrg[data$wloss_mrg == 1], "Contact Age by Weight Loss (Yes)")
violin_contact_age(data$snml_vage_mrg[data$wloss_mrg == 2], "Contact Age by Weight Loss (No)")
violin_contact_age(data$snml_vage_mrg[data$wloss_mrg == 3], "Contact Age by Weight Loss (Don't Know)")
violin_contact_age(data$snml_vage_mrg[is.na(data$wloss_mrg)], "Contact Age by Weight Loss (NA)")

violin_contact_age(data$snml_vage_mrg[data$cough_mrg == 1], "Contact Age by Cough (Yes)")
violin_contact_age(data$snml_vage_mrg[data$cough_mrg == 2], "Contact Age by Cough (No)")
violin_contact_age(data$snml_vage_mrg[data$cough_mrg == 3], "Contact Age by Cough (Don't Know)")
violin_contact_age(data$snml_vage_mrg[is.na(data$cough_mrg)], "Contact Age by Cough (NA)")

violin_contact_age(data$snml_vage_mrg[data$smoke_mrg == 1], "Contact Age by Smoking (Yes)")
violin_contact_age(data$snml_vage_mrg[data$smoke_mrg == 2], "Contact Age by Smoking (No)")
violin_contact_age(data$snml_vage_mrg[data$smoke_mrg == 3], "Contact Age by Smoking (Don't Know)")
violin_contact_age(data$snml_vage_mrg[is.na(data$smoke_mrg)], "Contact Age by Smoking (NA)")

violin_contact_age(data$snml_vage_mrg[data$drug_mrg == 1], "Contact Age by Drug Use (Yes)")
violin_contact_age(data$snml_vage_mrg[data$drug_mrg == 2], "Contact Age by Drug Use (No)")
violin_contact_age(data$snml_vage_mrg[data$drug_mrg == 3], "Contact Age by Drug Use (Don't Know)")
violin_contact_age(data$snml_vage_mrg[is.na(data$drug_mrg)], "Contact Age by Drug Use (NA)")
```



### **Statistical Tests**

**Statistical tests of network differences between cHEU and cHUU**

- Differences in average/median network size (ttest or Wilcoxon)
- Differences in average/median network density (ttest or Wilcoxon)
- Differences in proportion of adult men [network heterogeneity] (ttest or Wilcoxon)
- Differences in proportion with TB history [network heterogeneity] (ttest or Wilcoxon)
    - Within Household A
    - Outside of Household A


```{r}
household_density <- function(hh){ # simpson concentration index (SCI) 
    hh <- hh[!is.na(hh)]
    n = length(hh)
    if (n<2){
        return(NA_real_)
    }
    hh_table = table(hh)
    p = as.numeric(hh_table) / sum(hh_table) 
    sum(p^2) # probability that any two randomly sampled points in the ego are from the same household
}

df <- data %>%
    mutate(group = case_when(
        cherish_sn_study_group_mrg == 1 ~ "CHEU",
        cherish_sn_study_group_mrg == 2 ~ "CHUU",
        TRUE ~ NA_character_
    ),
    adult_male = (snml_sex_mrg == 1) & (snml_vage_mrg %in% c(6,7)),
    previos_tb = (tb_dx_mrg == 1)
    )
# head(df)

ego_stats_df <- df %>%
    filter(!is.na(group)) %>%
    group_by(record_id, group) %>%
    summarise(
        # this is the density metrics and the proportion of adult males and previous tb 
        n_contacts = n_distinct(contact_id),
        density = household_density(hh_id_unified_),
        prop_adult_men = mean(adult_male, na.rm = T),
        prop_prev_tb = mean(previos_tb, na.rm = T),

        # in household A
        n_a = n_distinct(contact_id[hh_id_unified_ == 1]),
        prop_adult_men_a = mean(adult_male[hh_id_unified_ == 1], na.rm = T),
        prop_prev_tb_a = mean(previos_tb[hh_id_unified_ == 1], na.rm = T),

        # not in household A
        n_nota = n_distinct(contact_id[!is.na(hh_id_unified_) & hh_id_unified_ != 1]),
        density_nota = household_density(hh_id_unified_[!is.na(hh_id_unified_) & hh_id_unified_ != 1]),
        prop_adult_men_nota = mean(adult_male[!is.na(hh_id_unified_) & hh_id_unified_ != 1], na.rm = T),
        prop_prev_tb_nota = mean(previos_tb[!is.na(hh_id_unified_) & hh_id_unified_ != 1], na.rm = T),
        .groups = "drop"
    )

head(ego_stats_df)
dim(ego_stats_df)
```



```{r}
tests <- c("n_contacts","density","prop_adult_men","prop_prev_tb",
           "n_a","prop_adult_men_a","prop_prev_tb_a",
           "n_nota","density_nota","prop_adult_men_nota","prop_prev_tb_nota")

results = do.call(rbind, lapply(tests, function(v){
    wt <- wilcox.test(ego_stats_df[[v]] ~ ego_stats_df$group,
                    exact = FALSE, conf.int = TRUE)
    data.frame(
        metric = v,
        W = unname(wt$statistic),
        p = wt$p.value,
        HL = if (!is.null(wt$estimate)) unname(wt$estimate) else NA_real_,
        CI_low = if (!is.null(wt$conf.int)) unname(wt$conf.int[1]) else NA_real_,
        CI_high= if (!is.null(wt$conf.int)) unname(wt$conf.int[2]) else NA_real_,
        stringsAsFactors = FALSE
        )
}))

results
```


**Statistical tests of network differences between 0yo, 1yo, and 2yo**

- Differences in average/median network size (ttest or Wilcoxon)
- Differences in average/median network density (ttest or Wilcoxon)
- Differences in proportion of adult men [network heterogeneity] (ttest or Wilcoxon)
- Differences in proportion with TB history [network heterogeneity] (ttest or Wilcoxon)
    - Within Household A
    - Outside of Household A


```{r}
ego_age <- df %>%
  distinct(record_id, enrol_age_yrs_mrg) %>%
  filter(enrol_age_yrs_mrg %in% c(0,1,2)) %>%
  mutate(index_age = factor(enrol_age_yrs_mrg, levels = c(0,1,2),
                            labels = c("0yo","1yo","2yo")))

ego_stats_age <- ego_stats_df %>%
  left_join(ego_age %>% dplyr::select(record_id, index_age), by = "record_id") %>%
  filter(!is.na(index_age))

vars <- c("n_contacts","density","prop_adult_men","prop_prev_tb",
          "n_a","prop_adult_men_a","prop_prev_tb_a",
          "n_nota","density_nota","prop_adult_men_nota","prop_prev_tb_nota")

res_kw <- do.call(rbind, lapply(vars, function(v){
  kt <- kruskal.test(ego_stats_age[[v]] ~ ego_stats_age$index_age)  # index_age must be 0yo/1yo/2yo factor
  data.frame(
    metric = v,
    chi_sq = unname(kt$statistic),
    df = unname(kt$parameter),
    p = kt$p.value,
    stringsAsFactors = FALSE
  )
}))
res_kw
```



**Statistical tests of network differences between cHEU and cHUU, adjusted for age**
- Differences in average/median network size (adjusted negative binomial regression)
- Differences in average/median network density (adjusted linear regression)
- Differences in proportion of adult men [network heterogeneity] (adjusted linear regression)
- Differences in proportion with TB history [network heterogeneity] (adjusted linear regression)
    - Within Household A
    - Outside of Household A


```{r}
# differences in median network size 
library(MASS)
ego_stats_age$group <- relevel(factor(ego_stats_age$group), ref = "CHUU")
m_size <- glm.nb(n_contacts ~ group + index_age, data = ego_stats_age)
summary(m_size)
```





```{r}
exp(coef(m_size)["groupCHEU"])        
exp(confint(m_size)["groupCHEU"])   
```





```{r}
m_density <- lm(density ~ group + index_age, data = ego_stats_age)
summary(m_density)
```





```{r}
coef(summary(m_density))["groupCHEU", ]
confint(m_density)["groupCHEU", ]
```





```{r}
m_am <- lm(prop_adult_men ~ group + index_age, data = ego_stats_age)
summary(m_am)
coef(summary(m_am))["groupCHEU", ]
confint(m_am)["groupCHEU", ]
```



```{r}
m_tb <- lm(prop_prev_tb ~ group + index_age, data = ego_stats_age)
summary(m_tb)
coef(summary(m_tb))["groupCHEU", ]
confint(m_tb)["groupCHEU", ]
```






```{r}
m_tb_a <- lm(prop_prev_tb_a ~ group + index_age, data = ego_stats_age)
summary(m_tb_a)
coef(summary(m_tb_a))["groupCHEU", ]
confint(m_tb_a)["groupCHEU", ]
```





```{r}
m_tb_out <- lm(prop_prev_tb_nota ~ group + index_age, data = ego_stats_age)
summary(m_tb_out)
coef(summary(m_tb_out))["groupCHEU", ]
confint(m_tb_out)["groupCHEU", ]
```



### **Wilcox Boxplots**

```{r}
pkgs <- c("ggpubr")
to_install <- pkgs[!vapply(pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(to_install)) install.packages(to_install)
library(ggpubr)

# inputs you already have:
# ego_stats_age (one row per index) with columns index_age + vars
# res_kw with columns metric + p

vars <- c("n_contacts","density","prop_adult_men","prop_prev_tb",
          "n_a","prop_adult_men_a","prop_prev_tb_a",
          "n_nota","density_nota","prop_adult_men_nota","prop_prev_tb_nota")

ego_stats_age$index_age <- factor(ego_stats_age$index_age, levels = c("0yo","1yo","2yo"))
comps <- list(c("0yo","1yo"), c("0yo","2yo"), c("1yo","2yo"))

make_one <- function(v){
  kw_p <- res_kw$p[match(v, res_kw$metric)]  # pull KW p from your dataframe

  ggboxplot(
    ego_stats_age, x = "index_age", y = v,
    add = "jitter", add.params = list(alpha = 0.25, size = 1),
    outlier.alpha = 0.25
  ) +
    labs(
      title = v,
      subtitle = paste0("KW p=", format.pval(kw_p, digits = 2)),
      x = NULL, y = NULL
    ) +
    stat_compare_means(
      comparisons = comps,
      method = "wilcox.test",
      label = "p.signif",                 # asterisks
      method.args = list(exact = FALSE),  # avoids ties warnings
      hide.ns = T,                     # change to FALSE if you want ns shown
      tip.length = 0.01,
      size = 4,
      vjust = 1
    ) +
    theme_bw(base_size = 12) +
    theme(
      plot.title = element_text(size = 11, face = "bold"),
      plot.subtitle = element_text(size = 9)
    )
}

plots <- lapply(vars, make_one)

# 4 x 3 grid
G <- ggarrange(plotlist = plots, ncol = 4, nrow = 3, align = "hv")
G
```

